Unicode Implementation Issues \thanks { This is a working document for the Rexx Language Association (RexxLA). There will be one or more separate requirements documents. It is written in \LaTeX\ \fmtversion\ and was rendered \today\ using \enginedetails. }

### 2024-03-23 ARB/standards/work-in-progress/unicode/Symposium 2024 review/reply-to-jlf-review.md

# Reply to Jean Louis Faucher review of the TUTOR article

I am replying in markdown format, because pdf is not very useable to have a dialog.

Please note that the version you have reviewed is not final. I have uploaded the final, published, versions of the articles in the Unicode/publications/35 subdirectory.

I will not copy some of your comments in the pdf, and intersperse my own comments in between. Page references are to the version you reviewed, not to the final document.

(p. 33)
>Inheritance from String.
>
>Each time a new instance of Byte, Codepoints, Graphemes or Text is created, a new copy of the original string is created. That can be confirmed by displaying the identityhash of the ~string.
>
>Not a problem for a prototype, but would be a problem for a real implementation. As written elsewhere, these 4 instances are a view of the same string. So it would be better to share this same string between the 4 views.
>
>Executor provides only Text, which correspond roughly to Tutor’s Graphemes (no normalization, but also no automatic conversion to UTF-8).  The original string is referenced (delegation), not inherited.
>That doesnt forbid to offer the same interface as .String.

OK. As you say, TUTOR is a prototype; I have not strived for efficiency. I am also under the impression that the code is much simpler if I use inheritance instead of delegation (but I might be wrong).

(p. 34)
>Executor applies NFC normalization on the fly when strict comparison of text.
>
>The difference between strict and non-strict is the management of blanks and numeric values.
>
>Rexxref: The two strings must be identical (character by character) and of the same length to be considered strictly equal. 
>No padding, no attempt to perform a numeric comparison

My interpretation of strict equality forbids any normalization. I understand "character by character" as _binary_ equality. 
If you normalize on the fly, you lose a first-class operator to determine whether two strings are binary identical.

My impression is that normalization pertains to _non-strict_ equality. "Normal" equality, that is, non-strict equality, is, somehow, the default
one: blanks are ignored, numbers have to be equivalent, etc., ... and "e" + acute accent is the same as "é".

OTOH, "==", I think, should mean "identical", that is, binary identical.

### 2024-03-23 ARB/standards/work-in-progress/unicode/Symposium 2024 review/reply-to-jlf-review.md

My interpretation of strict equality forbids any normalization. I understand "character by character" as _binary_ equality. 
If you normalize on the fly, you lose a first-class operator to determine whether two strings are binary identical.

My impression is that normalization pertains to _non-strict_ equality. "Normal" equality, that is, non-strict equality, is, somehow, the default
one: blanks are ignored, numbers have to be equivalent, etc., ... and "e" + acute accent is the same as "é".

OTOH, "==", I think, should mean "identical", that is, binary identical.

(p. 43)
>Can’t resist to remind my point of view, but feel free to ignore, was already discussed at one moment and written somewhere in the draft documents.
>
>It’s well defined: it’s what the interpreter exposes to you via the native C2X.
>
>Netrexx: UCS-2 (16 bits)
>
>Crexx: UTF-8
>
>all other interpreters: byte, whatever the encoding.
>
>You defined several views. Just apply these views on the native C2X. That will be exactly the same hex digits than the native C2X, but grouped according the view.

I think the final version refers to it (not sure about the version you have commented): it appears that recent versions of Java store Unicode strings as ISO-8859-1 when all 
codepoints are < "FF"X. This means that the notion of "internal representation" does no longer make sense (or varies according to whether a string contains a single
codepoint >= "100"X, which is ridiculous), as it is the remnant of a situation which is no longer true: when strings were manageable by a simple adlen pair, and were
composed of bytes.

(Ibid.)
>Hum… Don’t you have another BIF that convert to  a target encoding?

Yes. The addition of a second, encoding parameter to C2X is a compatibility aid: if you want NetRexx behaviour, for example, you just add "UTF-16" and you are done.

(p. 45)
> This is not very usual to return 2 results at once. Sounds like an internal optimization for your implementation.
> 
> UTF8()) returns a stem.

You are right, there are several details in the current version that are internal optimization leftovers. Will try to fix in a subsequent release.

### 2024-03-23 ARB/standards/work-in-progress/unicode/Symposium 2024 review/reply-to-jlf-review.md

(Ibid.)
>Hum… Don’t you have another BIF that convert to  a target encoding?

Yes. The addition of a second, encoding parameter to C2X is a compatibility aid: if you want NetRexx behaviour, for example, you just add "UTF-16" and you are done.

(p. 45)
> This is not very usual to return 2 results at once. Sounds like an internal optimization for your implementation.
> 
> UTF8()) returns a stem.

You are right, there are several details in the current version that are internal optimization leftovers. Will try to fix in a subsequent release. https://github.com/RexxLA/rexx-repository/issues/10

(Ibid.)
>This decode/encode reminds me Python.
>
>But your functions seems to be still a draft…
>
>with decode, you can transcode from a source encoding to a target encoding, where the target encoding is limited to UTF-8/16/32?

Only -8 and -32. This is another leftover. I need the -32 version internally, but I should most probably not expose that as an API.

(cont.)
>encode is also doing a transcoding… Only the target encoding is specified.
>The source encoding is UTF-8/16/32? Why is it not needed to specify it here?

My way of understanding the standard is the following: Unicode is the _lingua franca_ of all encodings. You only need a DECODE function that
specifies an encoding E (this converts from E to Unicode, our _lingua franca_), and a ENCODE function: this converts from Unicode,
our _lingua franca_, to a specified encoding E. You do not need a source encoding for ENCODE, since your source string is always (abstract)
Unicode.

If you want to transcode, say, from UTF-32 to another encoding E, you should first DECODE from UTF-32, and then ENCODE to E.

Most probably, the fact that DECODE allows a limited set of target Unicode encodings is misleading, and therefore a bad design decision. Thanks for bringing
this aspect to my attention.

### 2024-03-23 ARB/standards/work-in-progress/unicode/Symposium 2024 review/reply-to-jlf-review.md

My way of understanding the standard is the following: Unicode is the _lingua franca_ of all encodings. You only need a DECODE function that
specifies an encoding E (this converts from E to Unicode, our _lingua franca_), and a ENCODE function: this converts from Unicode,
our _lingua franca_, to a specified encoding E. You do not need a source encoding for ENCODE, since your source string is always (abstract)
Unicode.

If you want to transcode, say, from UTF-32 to another encoding E, you should first DECODE from UTF-32, and then ENCODE to E.

Most probably, the fact that DECODE allows a limited set of target Unicode encodings is misleading, and therefore a bad design decision. Thanks for bringing
this aspect to my attention.

(p. 47)
>I’m not aware of WTF32, and I don’t see how it could be needed. There is no low/high surrogates in UTF32.
>Do you have links?

I can't seem to be able to recover the links I once had, my bad; I have not invented the term. Just because UTF-32 does not allow
lone surrogates, we _need_ WTF-32. For example, when handling WTF-8, we may find lone surrogates. Either we forbid transforming
WTF-8 to a four-bytes per codepoint format (bad), or we have to implement WTF-32.

(p. 48)
>U2C() is cited once elsewhere in this document. But not yet implemented? or discarded?

Forgot to document it :/ It was implemented as a method, but not as a (new) BIF. Fixed: https://github.com/RexxLA/rexx-repository/blob/master/ARB/standards/work-in-progress/unicode/UnicodeTools/doc/new-functions.md#u2c-unicode-to-character

### 2024-03-23 ARB/standards/work-in-progress/unicode/Symposium 2024 review/reply-to-jlf-review.md

I can't seem to be able to recover the links I once had, my bad; I have not invented the term. Just because UTF-32 does not allow
lone surrogates, we _need_ WTF-32. For example, when handling WTF-8, we may find lone surrogates. Either we forbid transforming
WTF-8 to a four-bytes per codepoint format (bad), or we have to implement WTF-32.

(p. 48)
>U2C() is cited once elsewhere in this document. But not yet implemented? or discarded?

Forgot to document it :/ It was implemented as a method, but not as a (new) BIF. Fixed: https://github.com/RexxLA/rexx-repository/blob/master/ARB/standards/work-in-progress/unicode/UnicodeTools/doc/new-functions.md#u2c-unicode-to-character

(Ibid.)
>Should I provide U2C? it would support only the U+ notation.
>Yes, having C2U/U2C would be consistent with C2X/X2C.
>The target encoding would be the default encoding.

I can see that this is a note to yourself. Just wanted to add that this is just the idea: consistency with C2X/X2C. My U2C allows names, aliases and labels, in addition to codepoints and U+codepoints.

(Ibid.)
>Low level, so maybe you need it for your implementation.
>
>But from an end-user perspective, don’t you have a BIF that returns all the properties of a unicode character? including its name and aliases. Yes: UNICODE()

I think that in many cases it may be very convenient to specify a character by name. Not so much low-level, I think.

The idea of subsuming this in UNICODE is great. https://github.com/RexxLA/rexx-repository/issues/11

### 2024-03-23 ARB/standards/work-in-progress/unicode/Symposium 2024 review/reply-to-jlf-review.md

(Ibid.)
>Low level, so maybe you need it for your implementation.
>
>But from an end-user perspective, don’t you have a BIF that returns all the properties of a unicode character? including its name and aliases. Yes: UNICODE()

I think that in many cases it may be very convenient to specify a character by name. Not so much low-level, I think.

The idea of subsuming this in UNICODE is great. https://github.com/RexxLA/rexx-repository/issues/11

(p. 50)
>You could also support a Unicode character name, but I understand that it collides with the other form UNICODE(string…).

Again a great idea. https://github.com/RexxLA/rexx-repository/issues/12

(p. 55)
>Under macOS, and probably under Linux:
>
>  rlwrap rexx rxutry
>
>to have an history of commands.

Thanks for the tip! I will document it in the following release.

(p. 56)
>Something specific to ooRexx (and maybe Netrexx?): the caseless methods.

Have not delved into case folding still. Thanks for the info!

### 2024-03-23 ARB/standards/work-in-progress/unicode/Symposium 2024 review/reply-to-jlf-review.md

# Reply to Jean Louis Faucher review of the TUTOR article

I am replying in markdown format, because pdf is not very useable to have a dialog.

Please note that the version you have reviewed is not the final one. I have uploaded the final, published, versions of the articles in the Unicode/publications/35 subdirectory.

I will now copy some of your comments in the pdf, and intersperse my own comments in between. Page references are to the version you reviewed, not to the final document.

### 2024-03-23 ARB/standards/work-in-progress/unicode/Symposium 2024 review/reply-to-jlf-review.md

(p. 34)
>Executor applies NFC normalization on the fly when strict comparison of text.
>
>The difference between strict and non-strict is the management of blanks and numeric values.
>
>Rexxref: The two strings must be identical (character by character) and of the same length to be considered strictly equal. 
>No padding, no attempt to perform a numeric comparison

My interpretation of strict equality forbids any normalization. I understand "character by character" as _binary_ equality. 
If you normalize on the fly, you lose a first-class operator to determine whether two strings are binary identical.

My impression is that normalization pertains to _non-strict_ equality. "Normal" equality, that is, non-strict equality, is, somehow, the default
one: blanks are ignored, numbers have to be equivalent, etc., ... and "e" + acute accent is the same as "é".

"==", I think, should mean "identical", that is, binary identical.

@@

(p. 43)
>Can’t resist to remind my point of view, but feel free to ignore, was already discussed at one moment and written somewhere in the draft documents.
>
>It’s well defined: it’s what the interpreter exposes to you via the native C2X.
>
>Netrexx: UCS-2 (16 bits)
>
>Crexx: UTF-8
>
>all other interpreters: byte, whatever the encoding.
>
>You defined several views. Just apply these views on the native C2X. That will be exactly the same hex digits than the native C2X, but grouped according the view.

I think the final version refers to it (not sure about the version you have commented): it appears that recent versions of Java store Unicode strings as ISO-8859-1 when all 
codepoints are < "FF"X. This means that the notion of "internal representation" does no longer make sense (or varies according to whether a string contains a single
codepoint >= "100"X, which is ridiculous), as it is the remnant of a situation which is no longer true: when strings were manageable by a simple adlen pair and a chunk of bytes.

### 2024-03-23 ARB/standards/work-in-progress/unicode/Symposium 2024 review/reply-to-jlf-review.md

(Ibid.)
>This decode/encode reminds me Python.
>
>But your functions seems to be still a draft…
>
>with decode, you can transcode from a source encoding to a target encoding, where the target encoding is limited to UTF-8/16/32?

Only -8 and -32. This is another leftover. I need the -32 version internally, but I should most probably not expose that as an API.

(cont.)
>encode is also doing a transcoding… Only the target encoding is specified.
>The source encoding is UTF-8/16/32? Why is it not needed to specify it here?

My way of understanding the standard is the following: Unicode is the _lingua franca_ of all encodings. You only need a DECODE function that
specifies an encoding E (this converts from E to Unicode, our _lingua franca_), and an ENCODE function: this converts from Unicode,
our _lingua franca_, to a specified encoding E. You do not need a source encoding for ENCODE, since your source string is always (abstract)
Unicode.

### 2024-04-12 ARB/standards/work-in-progress/document/tex/standard/built-in_classes.tex
add filenames

::class object

::method new class Returns a new instance of the receiver class.

call Config ObjectNew return \#Outcome

::method `==' `==' with no argument gives a hash value in OOI.

call Config ObjectCompare \#Receiver, \#Arg.1

if \#Outcome == `equal' then return `1'

else return `0'

::method `\textless\textgreater!'

use arg a

return \self==a

::method `\textgreater\textless{}' forward message
`\textless\textgreater{}'

::method `=' forward message `=='

::method `=! forward message'\textless\textgreater'

::method `==' forward message `\textless\textgreater{}'

::method copy Returns a copy of the receiver object. The copied object
has the same methods as the receiver object and an equivalent set of
object variables, with the same values.

call Config ObjectCopy \#Receiver

return \#Outcome

Since we have var\_empty we could save a primitive by rendering `new' as
`copy' plus `empty'. ::method defaultname

@@

call var\_value \#Receiver, `\#Human', `0'

return \#Outcome

This field would have been filled in at `NEW' time.

::method `OBJECTNAMES' /* rvSTRING */

@@

call var\_set \#Receiver, \#ObjectName, `0', \#Arg.1 return

Initialized to \#Human? Or ObjectName does forwarding until assigned to?

::method objectname

Returns the receiver object's name (which the OBJECTNAME= method sets).
call var\_value \#Receiver, \#ObjectName, `0' return \#Outcome

::method string

Returns a human-readable string representation for the object. return
\#Receiver\textasciitilde ObjectName

::method class Returns the class object that received the message that
created the object.

call var value \#Receiver, \#IsA, `0' return \#Outcome

::method setmethod /* rSTRING oSTRING/METHOD/ARRAY */ Adds a method to
the receiver object's collection of object methods.

Is `object methods' what is intended; you add to a class without adding
to its instance methods? Yes. if \#ArgExists.2 then m = Arg.2

else m = .NIL call set\_var \#Receiver, `METHODS.'\#Arg.1, `1', m return
::method hasmethod /* rSTRING */

@@

This presumably means inherited as well as SETMETHOD ones. What about
ones set to .NIL?

Need to use the same search as for sending.

::method unsetmethod private

Removes a method from the receiver object's collection of object
methods. Use var\_drop

Private means Receiver = Self check.

::method request /* rSTRING */ Returns an object of the specified class,
or the NIL object if the request cannot be satisfied. t = `MAKE'\#Arg.1

if \#Receiver\textasciitilde hasmethod(t) then return .NIL forward
message(t) array ()

::method run private /* rMETHOD Ugh keyoptions */ Runs the specified
method. The method has access to the object variables of the receiver
object, just as if the receiver object had defined the method by using
SETMETHOD.

::smethod startat Undocumented?

::method start /* rMESSAGE oArglist */

Returns a message object and sends it a START message to start
concurrent processing. ::method init

@@

::class class

Lots of these methods are both class and instance. \textbar{} don't know
whether to list them twice. ::method new class /* OARGLIST */

Returns a new instance of the receiver class, whose object methods are
the instance methods of the class. This method initializes a new
instance by running its INIT methods.

::method subclass class

@@

::method subclasses class

@@

::method define class /* rSTRING oMETHOD */

@@

::method delete

@@

Builtin classes cannot be altered.

::method method class /* rSTRING */

@@

Do we have to keep saying ``method object'' as opposed to ``method''
because ``method name'' exists? ::method querymixinclass

@@

::method mixinclass class /* 3 of em */

@@

::method inherit class /* rCLASS oCLASS */

Causes the receiver class to inherit the instance and class methods of
the class object specified. The

optional class is a class object that specifies the position of the new
superclass in the list of superclasses. ::method uninherit class /*
rCLASSOBJ */

Nullifies the effect of any previous INHERIT message sent to the
receiver for the class specified. ::smethod enhanced class /*
rCOLLECTION oArgs */

Returns an enhanced new instance of the receiver class, with object
methods that are the instance

methods of the class enhanced by the methods in the specified collection
of methods. ::method baseclass class

Returns the base class associated with the class. If the class is a
mixin class, the base class is the first superclass that is not also a
mixin class. If the class is not a mixin class, then the base class is
the class receiving the BASECLASS message.

::method superclasses class Returns the immediate superclasses of the
receiver class in the form of a single-index array of the

required size. ::method id class

Returns a string that is the class identity (instance SUBCLASS and
MIXINCLASS methods.)

::method metaclass class

Returns the receiver class's default metaclass. ::method methods class
/* oCLASSOBJECT */

Returns a supplier object for all the instance methods of the receiver
class and its superclasses, if no argument is specified. \#\#\# The
method class

::class method

::method new class /* rSTRING rSOURCE */ Returns a new instance of
method class, which is an executable representation of the code
contained in

the source. ::method setprivate

Specifies that a method is a private method. ::method setprotected

::method setsecuritymanager

::method setguarded Reverses any previous SETUNGUARDED messages,
restoring the receiver to the default guarded

status. ::method setunguarded

Lets an object run a method even when another method is active on the
same object. If a method object

does not receive a SETUNGUARDED message, it requires exclusive use of
its object variable pool. ::method source

Returns the method source code as a single index array of source lines.
::method interface

::method setinterface

11.3 The string class The string class provides conventional strings and
numbers.

Some differences from REXX class of NetRexx. ::class string

::method new class

::method `\textquotesingle{} We can do all the operators by appeal to
classic section 7. ::method'-' ::method `-'

use arg a return \a General problem of making the error message come
right.

:method :method :method :method :method :method :method :method :method
:method :method :method :method :method :method :method :method :method
:method :method :method :method :method :method :method :method :method
:method :method

:method :method :method :method :method :method :method :method :method
:method :method :method :method :method :method :method :method :method
:method :method :method :method :method :method :method :method :method
:method :method :method :method :method :method :method :method :method
:method :method :method :method :method

All

centre center changestr compare copies counstr datatype delstr delword
insert lastpos left length overlay pos reverse right space strip substr
subword translate verify word wordindex wordlength wordpos words abs
format max

min

sign trunc B2x bitand bitor bitxor C2D

C2x

D2xX ::smethod D2C ::smethod X2B ::smethod X2C ::smethod X2D ::method
string

::method makestring \#\#\# The array class

@@

To be done. Dimensionality set at first use. Count commas, not classic
arg().

::class array

::method new class /* 0 or more WHOLE\textgreater=0 */

@@

::method of class /* 0 or more ANY */

Returns a newly created single-index array containing the specified
value objects. ::method put /* rANY one or more WHOLE\textgreater0 */

Makes the object value a member item of the array and associates it with
the specified index or indexes. ::method ' {[}{]}=' /* 1 or more
WHOLE\textgreater0O */

This method is the same as the PUT method.

::method at /* 1 or more WHOLE\textgreater0O */

@@

::method `{[}{]}' /* 1 or more WHOLE\textgreater0O */

@@

::method remove /* 1 or more WHOLE\textgreater0O */

Returns and removes the member item with the specified index or indexes
from the array. ::method hasindex /* 1 or more WHOLE\textgreater0O */

@@

::method items /* (None) \emph{/ Returns the number of items in the
collection. ::method dimension /} OWHOLE\textgreater0O */

@@

::method size /* (None) */

@@

::method first /* (None) */

Returns the index of the first item in the array, or the NIL object if
the array is empty.

::method last /* (None) */

Returns the index of the last item in the array, or the NIL object if
the array is empty.

::method next /* rcWHOLE\textgreater O */

Returns the index of the item that follows the array item having the
specified index or returns the NIL object if the item having that index
is last in the array.

::method previous /* rcWHOLE\textgreater O */

Returns the index of the item that precedes the array item having index
index or the NIL object if the item having that index is first in the
array.

::method makearray /* (None) */

@@

Returns a new array (of the same class as the receiver) containing
selected items from the receiver array. The first item in the new array
is the item corresponding to index start (the first argument) in the
receiver

array. ::method supplier /* (None) \emph{/ Returns a supplier object for
the collection. ::method section /} rcWHOLE\textgreater0O
oOWHOLE\textgreater=0 */

\section{The supplier class}\label{the-supplier-class}

A supplier object enumerates the items a collection contained at the
time of the supplier's creation. ::class supplier

::method new class /* rANYARRAY rINDEXARRAY */ Returns a new supplier
object.

::method index

Returns the index of the current item in the collection. ::method next

@@

::method item

@@

::method available

Returns 1 (true) if an item is available from the supplier (that is, if
the ITEM method would return a value). Returns 0 (false) otherwise.

11.5 The message class ::class message

::method init class /* Ugh */

@@

::method completed

Returns 1 if the message object has completed its message; returns 0
otherwise. ::method notify /* rMESSAGE */

Requests notification about the completion of processing for the message
SEND or START sends. ::method start /* oANY */

Sends the message for processing concurrently with continued processing
of the sender. ::method send /* oANY */

@@

::method result

Returns the result of the message SEND or START sends.

>>>>

\lstinputlisting[language=rexx,label=object.rexx,caption=object.rexx]{object.rexx}

Returns a new instance of the receiver class.

\lstinputlisting[language=rexx,label=config.rexx,caption=config.rexx]{config.rexx}

\emph{\texttt{\textquotesingle{}==\textquotesingle{}} with no argument
gives a hash value in OOI.}

\lstinputlisting[language=rexx,label=doubleequals.rexx,caption=doubleequals.rexx]{doubleequals.rexx}

Returns a copy of the receiver object. The copied object has the same
methods as the receiver object and an equivalent set of object
variables, with the same values.

\lstinputlisting[language=rexx,label=objectcopy.rexx,caption=objectcopy.rexx]{objectcopy.rexx}

\emph{Since we have \texttt{var\_empty} we could save a primitive by
rendering `new' as `copy' plus `empty'.}

\lstinputlisting[language=rexx,label=defaultname.rexx,caption=defaultname.rexx]{defaultname.rexx}

@@

\lstinputlisting[language=rexx,label=receiver.rexx,caption=receiver.rexx]{receiver.rexx}

\emph{This field would have been filled in at `NEW' time.}

\lstinputlisting[language=rexx,label=objectnames.rexx,caption=objectnames.rexx]{objectnames.rexx}

@@

\lstinputlisting[language=rexx,label=varset.rexx,caption=varset.rexx]{varset.rexx}

\emph{Initialized to \#Human? Or ObjectName does forwarding until
assigned to?}

\lstinputlisting[language=rexx,label=objectname,caption=objectname]{objectname}

Returns the receiver object's name (which the \texttt{OBJECTNAME=}
method sets).

\lstinputlisting[language=rexx,label=receiverObjectName.rexx,caption=receiverObjectName.rexx]{receiverObjectName.rexx}

Returns a human-readable string representation for the object.

\lstinputlisting[language=rexx,label=methodclass.rexx,caption=methodclass.rexx]{methodclass.rexx}

Returns the class object that received the message that created the
object.

\lstinputlisting[language=rexx,label=setmethod.rexx,caption=setmethod.rexx]{setmethod.rexx}

Adds a method to the receiver object's collection of object methods.

\emph{Is `object methods' what is intended; you add to a class without
adding to its instance methods? Yes.}

\lstinputlisting[language=rexx,label=hasmethod.rexx,caption=hasmethod.rexx]{hasmethod.rexx}

@@

\emph{This presumably means inherited as well as \texttt{SETMETHOD}
ones. What about ones set to \texttt{.NIL}?}

\emph{Need to use the same search as for sending.}

\lstinputlisting[language=rexx,label=unsetmethod.rexx,caption=unsetmethod.rexx]{unsetmethod.rexx}

Removes a method from the receiver object's collection of object
methods.

\emph{Use \texttt{var\_drop}}

\emph{Private means Receiver = Self check.}

\lstinputlisting[language=rexx,label=requestmethod.rexx,caption=requestmethod.rexx]{requestmethod.rexx}

Returns an object of the specified class, or the \texttt{NIL} object if
the request cannot be satisfied.

\lstinputlisting[language=rexx,label=runmethod.rexx,caption=runmethod.rexx]{runmethod.rexx}

Runs the specified method. The method has access to the object variables
of the receiver object, just as if the receiver object had defined the
method by using \texttt{SETMETHOD}.

\lstinputlisting[language=rexx,label=startat.rexx,caption=startat.rexx]{startat.rexx}

Returns a message object and sends it a \texttt{START} message to start
concurrent processing.

\lstinputlisting[language=rexx,label=init.rexx,caption=init.rexx]{init.rexx}

@@

\lstinputlisting[language=rexx,label=classclass.rexx,caption=classclass.rexx]{classclass.rexx}

\emph{Lots of these methods are both class and instance. I don't know
whether to list them twice.}

\lstinputlisting[language=rexx,label=newclass.rexx,caption=newclass.rexx]{newclass.rexx}

Returns a new instance of the receiver class, whose object methods are
the instance methods of the class. This method initializes a new
instance by running its \texttt{INIT} methods.

\lstinputlisting[language=rexx,label=subclassclass.rexx,caption=subclassclass.rexx]{subclassclass.rexx}

@@

\lstinputlisting[language=rexx,label=immediatesubclass,caption=immediatesubclass]{immediatesubclass}

@@

\lstinputlisting[language=rexx,label=defineclass.rexx,caption=defineclass.rexx]{defineclass.rexx}

@@

\lstinputlisting[language=rexx,label=deleteclass.rexx,caption=deleteclass.rexx]{deleteclass.rexx}

@@

\emph{Builtin classes cannot be altered.}

\lstinputlisting[language=rexx,label=methodmethod.rexx,caption=methodmethod.rexx]{methodmethod.rexx}

@@

\emph{Do we have to keep saying ``method object'' as opposed to
``method'' because ``method name'' exists?}

\lstinputlisting[language=rexx,label=querymixin.rexx,caption=querymixin.rexx]{querymixin.rexx}

@@

\lstinputlisting[language=rexx,label=newmixinclass.rexx,caption=newmixinclass.rexx]{newmixinclass.rexx}

@@

\lstinputlisting[language=rexx,label=inheritclass.rexx,caption=inheritclass.rexx]{inheritclass.rexx}

Causes the receiver class to inherit the instance and class methods of
the class object specified. The optional class is a class object that
specifies the position of the new superclass in the list of
superclasses.

\lstinputlisting[language=rexx,label=uninheritclass.rexx,caption=uninheritclass.rexx]{uninheritclass.rexx}

Nullifies the effect of any previous \texttt{INHERIT} message sent to
the receiver for the class specified.

\lstinputlisting[language=rexx,label=enhancedclass.rexx,caption=enhancedclass.rexx]{enhancedclass.rexx}

Returns an enhanced new instance of the receiver class, with object
methods that are the instance methods of the class enhanced by the
methods in the specified collection of methods.

\lstinputlisting[language=rexx,label=baseclassclass.rexx,caption=baseclassclass.rexx]{baseclassclass.rexx}

Returns the base class associated with the class. If the class is a
mixin class, the base class is the first superclass that is not also a
mixin class. If the class is not a mixin class, then the base class is
the class receiving the \texttt{BASECLASS} message.

\lstinputlisting[language=rexx,label=superclassclass.rexx,caption=superclassclass.rexx]{superclassclass.rexx}

Returns the immediate superclasses of the receiver class in the form of
a single-index array of the required size.

\lstinputlisting[language=rexx,label=methodid.rexx,caption=methodid.rexx]{methodid.rexx}

Returns a string that is the class identity (instance \texttt{SUBCLASS}
and \texttt{MIXINCLASS} methods.)

\lstinputlisting[language=rexx,label=metaclassclass.rexx,caption=metaclassclass.rexx]{metaclassclass.rexx}

Returns the receiver class's default metaclass.

\lstinputlisting[language=rexx,label=methodsclass.rexx,caption=methodsclass.rexx]{methodsclass.rexx}

Returns a supplier object for all the instance methods of the receiver
class and its superclasses, if no argument is specified.

\subsection{The method class}\label{the-method-class}

\lstinputlisting[language=rexx,label=methodclass.rexx,caption=methodclass.rexx]{methodclass.rexx}

Returns a new instance of method class, which is an executable
representation of the code contained in the source.

\lstinputlisting[language=rexx,label=setprivatemethod.rexx,caption=setprivatemethod.rexx]{setprivatemethod.rexx}

Specifies that a method is a private method.

\lstinputlisting[language=rexx,label=setvarious.rexx,caption=setvarious.rexx]{setvarious.rexx}

Reverses any previous \texttt{SETUNGUARDED} messages, restoring the
receiver to the default guarded status.

\lstinputlisting[language=rexx,label=setunguarded.rexx,caption=setunguarded.rexx]{setunguarded.rexx}

Lets an object run a method even when another method is active on the
same object. If a method object does not receive a \texttt{SETUNGUARDED}
message, it requires exclusive use of its object variable pool.

\lstinputlisting[language=rexx,label=methodsource.rexx,caption=methodsource.rexx]{methodsource.rexx}

Returns the method source code as a single index array of source lines.

\lstinputlisting[language=rexx,label=methodinterface,caption=methodinterface]{methodinterface}

\subsection{The string class}\label{the-string-class}

The string class provides conventional strings and numbers.

\emph{Some differences from REXX class of NetRexx.}

\lstinputlisting[language=rexx,label=stringclass.rexx,caption=stringclass.rexx]{stringclass.rexx}

\emph{We can do all the operators by appeal to classic section 7.}

\lstinputlisting[language=rexx,label=stringmethods,caption=stringmethods]{stringmethods}

\emph{General problem of making the error message come right.}

\lstinputlisting[language=rexx,label=stringoperatorsandmethods,caption=stringoperatorsandmethods]{stringoperatorsandmethods}

\subsection{The array class}\label{the-array-class}

@@

\emph{To be done. Dimensionality set at first use. Count commas, not
classic \texttt{arg()}.}

\lstinputlisting[language=rexx,label=arrayclass.rexx,caption=arrayclass.rexx]{arrayclass.rexx}

@@

\lstinputlisting[language=rexx,label=returnsingleindexarray.rexx,caption=returnsingleindexarray.rexx]{returnsingleindexarray.rexx}

Returns a newly created single-index array containing the specified
value objects.

\lstinputlisting[language=rexx,label=makearraymemberitem.rexx,caption=makearraymemberitem.rexx]{makearraymemberitem.rexx}

Makes the object value a member item of the array and associates it with
the specified index or indexes.

\lstinputlisting[language=rexx,label=operatorputmethodarray.rexx,caption=operatorputmethodarray.rexx]{operatorputmethodarray.rexx}

This method is the same as the \texttt{PUT} method.

\lstinputlisting[language=rexx,label=arraymethodat.rexx,caption=arraymethodat.rexx]{arraymethodat.rexx}

@@

\lstinputlisting[language=rexx,label=arraymethodatoperator.rexx,caption=arraymethodatoperator.rexx]{arraymethodatoperator.rexx}

@@

\lstinputlisting[language=rexx,label=arraymethodremove.rexx,caption=arraymethodremove.rexx]{arraymethodremove.rexx}

Returns and removes the member item with the specified index or indexes
from the array.

\lstinputlisting[language=rexx,label=arraymethodhasindex.rexx,caption=arraymethodhasindex.rexx]{arraymethodhasindex.rexx}

@@

\lstinputlisting[language=rexx,label=arraymethoditems.rexx,caption=arraymethoditems.rexx]{arraymethoditems.rexx}

Returns the number of items in the collection.

\lstinputlisting[language=rexx,label=arraymethoddimension.rexx,caption=arraymethoddimension.rexx]{arraymethoddimension.rexx}

@@

\lstinputlisting[language=rexx,label=arraymethodsize.rexx,caption=arraymethodsize.rexx]{arraymethodsize.rexx}

@@

\lstinputlisting[language=rexx,label=arraymethodfirst.rexx,caption=arraymethodfirst.rexx]{arraymethodfirst.rexx}

Returns the index of the first item in the array, or the \texttt{NIL}
object if the array is empty.

\lstinputlisting[language=rexx,label=arraymethodlast,caption=arraymethodlast]{arraymethodlast}

Returns the index of the last item in the array, or the \texttt{NIL}
object if the array is empty.

\lstinputlisting[language=rexx,label=arraymethodnext.rexx,caption=arraymethodnext.rexx]{arraymethodnext.rexx}

Returns the index of the item that follows the array item having the
specified index or returns the \texttt{NIL} object if the item having
that index is last in the array.

\lstinputlisting[language=rexx,label=arraymethodprevious.rexx,caption=arraymethodprevious.rexx]{arraymethodprevious.rexx}

Returns the index of the item that precedes the array item having index
index or the \texttt{NIL} object if the item having that index is first
in the array.

\lstinputlisting[language=rexx,label=arraymethodmakearray.rexx,caption=arraymethodmakearray.rexx]{arraymethodmakearray.rexx}

@@

Returns a new array (of the same class as the receiver) containing
selected items from the receiver array. The first item in the new array
is the item corresponding to index start (the first argument) in the
receiver array.

\lstinputlisting[language=rexx,label=arraymethodsupplier,caption=arraymethodsupplier]{arraymethodsupplier}

Returns a supplier object for the collection.

\lstinputlisting[language=rexx,label=arraymethodsection,caption=arraymethodsection]{arraymethodsection}

\subsection{The supplier class}\label{the-supplier-class}

A supplier object enumerates the items a collection contained at the
time of the supplier's creation.

\lstinputlisting[language=rexx,label=supplierclass.rexx,caption=supplierclass.rexx]{supplierclass.rexx}

Returns a new supplier object.

\lstinputlisting[language=rexx,label=suppliermethodindex.rexx,caption=suppliermethodindex.rexx]{suppliermethodindex.rexx}

Returns the index of the current item in the collection.

\lstinputlisting[language=rexx,label=suppliermethodnext.rexx,caption=suppliermethodnext.rexx]{suppliermethodnext.rexx}

@@

\lstinputlisting[language=rexx,label=suppliermethoditem.rexx,caption=suppliermethoditem.rexx]{suppliermethoditem.rexx}

@@

\lstinputlisting[language=rexx,label=supplierindexavailable.rexx,caption=supplierindexavailable.rexx]{supplierindexavailable.rexx}

Returns 1 (true) if an item is available from the supplier (that is, if
the \texttt{ITEM} method would return a value). Returns 0 (false)
otherwise.

\subsection{The message class}\label{the-message-class}

\lstinputlisting[language=rexx,label=messageclass.rexx,caption=messageclass.rexx]{messageclass.rexx}

@@

\lstinputlisting[language=rexx,label=messagecompletedmethod.rexx,caption=messagecompletedmethod.rexx]{messagecompletedmethod.rexx}

Returns 1 if the message object has completed its message; returns 0
otherwise.

\lstinputlisting[language=rexx,label=messagebotifymethod.rexx,caption=messagebotifymethod.rexx]{messagebotifymethod.rexx}

Requests notification about the completion of processing for the message
\texttt{SEND} or \texttt{START} sends.

\lstinputlisting[language=rexx,label=messagestartmethod.rexx,caption=messagestartmethod.rexx]{messagestartmethod.rexx}

Sends the message for processing concurrently with continued processing
of the sender.

\lstinputlisting[language=rexx,label=messagesendmethod.rexx,caption=messagesendmethod.rexx]{messagesendmethod.rexx}

@@

\lstinputlisting[language=rexx,label=methodresult.rexx,caption=methodresult.rexx]{methodresult.rexx}

Returns the result of the message \texttt{SEND} or \texttt{START} sends.
